### 2022-08-10 article.tex
Improve

Albeit hosted code is vastly superior over raw code
in case of reproduciblity and correctness,
there are some problems with hosted code as well.
The most important one is that hosted code needs maintanance:
the libraries that are used by an experiment
will change over time, inevitably causing the experiment to fail
over time. This process is called bit rot and is a fundamental

HIERO

problem \richel{ref}. A solution that has more 
longevity is to publishing the code installed in a virtual environment,
as described below.

A lesser problem, which is only a problem to some, is, that
the author of the code can be contacted and that bugs can be reported.
It can be argued that no energy should be wasted on published code.
This is however, a rather selfish point: a bug should be reported and
given proper attention, for the sake of the correctness of the 
results published.

@@

Here are multiple suggestions to make hosted code FAIR.
To make the hosted code findable, commonly a paper refers to its URL.
At the homepage of hosted code, there is no standardized metadata, to,
for example, link to the paper.
Using a popular code host ensures the code is easily accessible.
There is some interoperabily supplied by a version control manager,
such as in obtaining the history of the code.
The reusability of code is reasonably easy to assess, by using build badges.
However, there is no guarantee that a build badge that claimes success,
actually is an honest signal (i.e. that code may not actually build),
nor are there standards for what 'success' entails.

@@

The most reproducible way of submitting the code of an experiment,
is by providing the code and all it needs 
installed in a virtual container.
This is close to the golden standard suggested by 
\cite{peng2011reproducible} (see also Figure \ref{fig:peng2011reproducible}),
however, that was before virtualization existed.
Virtualization allows one to provide a standard environment,
that can be used on any other computer.
The best known examples of container software are Docker and Singularity,
where the latter is the only viable option to be able to run
an experiment on a high performance computing (HPC) environment,
such as a computer cluster.

@@

Unlike English, code is fragile in standing the test of time.
Containers can alleviate this, as these can be created at a point
in time and then run later. Because the container is stand-alone,
all libraries the code depends on remain in their current/old state.
As with hosted code, an experiment may not run with newer versions
of libraries anymore, but a containerized experiment can run regardless, if
one refrains from updating.

@@

There are some problems with providing a container with a runnable container.
First, a container is several gigabytes and need to be stored somewhere
online. Ideally, this would be at Syslabs, as that one is free, yet
there is a limit of 15 Gb. Also, Sylabs does not allow much metadata to
be entered, such as the URL of the article, nor the URL of hosted code.
Also, journals do not request hosted code, nor runnable code.

@@

The code that does the work in a computational experiment should
always be published together with a scholarly article.
For research to be reproducible, one ideally has access to
both the data used and the code.
In some fields, such as genetic epidemiology, the data is
sensitive, hence cannot be released,
yet there are methods being devised to run code on sensitive
data with assured privacy \cite{zhang2016review,azencott2018machine}.

Code has additional useful information, similar to confidence intervals,
that allow a reader to gauge how much he/she trusts the results.
The most important way to determine the quality of code
is the amount of unit tests.
When following a set of best practices, such as DevOps, TDD, Agile,
writing unit tests is an essential 
part in writing code.
The amount of unit tests is an honest signal 
for code correctness (i.e. it does what it is supposed to do, as opposed
to 'it does something').
In academia, to uncover the truth, code correctness is essential.
To make a comparison with cell biologists, where working
sterile is essential to have the correct results,
unit tests allow a computational biologist to have the same.

>>>>

Albeit hosted code is vastly superior over raw code
in case of reproduciblity and correctness,
there are some problems with hosted code as well.
The most important one is that hosted code needs maintanance:
the libraries that are used by an experiment
will change over time, inevitably causing the experiment to fail
over time. 
The degradation of software is a known feature for nearly 
four decades (see the definition of 'bit rot' at \cite{steele1983hacker}),
here I will use the more modern (and in this context superior)
definition of software collapse (\cite{hinsen2019dealing}),
in which software fails due to dependencies on other 
software.
Where this paper recommends using older software foundations 
with multiple implementations (\cite{hinsen2019dealing}),
there is a modern solution that potentially has more longevity:
to publish the code installed in a virtual environment
(including the dependencies used at that time), as described below.

There is another problem that authors of academic code face,
which is that they will be contacted with a user sending in a bug report.
One solution for the author is to ignore such emails
and it can be argued that no energy should be wasted on published code
and work on something new instead.
However, upon the find of a new bug, the question is:
does all the research that use that software still 
result in the same conclusions?
A code hosting website usually has a place to report bugs
that offer transparency in the handling of bug reports.

@@

Here are multiple suggestions to make hosted code FAIR.
To make the hosted code findable, commonly a paper refers to its URL.
At the homepage of hosted code (e.g. a GitHub repository), 
there is no standardized metadata, to,
for example, link to the paper.
Using a popular code host ensures the code is easily accessible.
There is some interoperabily supplied by a version control manager,
such as in obtaining the history of the code.
The reusability of code is reasonably easy to assess, 
by using so-called build badges.
However, there is no guarantee that a build badge that claimes success,
actually is an honest signal (i.e. that code may not actually build),
nor are there standards for what 'success' entails.

@@

The most reproducible way of submitting the code of an experiment,
is by providing the code with its dependencies installed in a virtual container.
This is close to the golden standard suggested by 
\cite{peng2011reproducible} (see also Figure \ref{fig:peng2011reproducible}),
however, that was before virtualization existed.
Virtualization allows one to provide a standard environment,
that can be used on any other computer.
The best known examples of container software are Docker and Singularity,
where the latter is the only viable option to be able to run
an experiment on a high performance computing (HPC) environment,
such as a computer cluster.

@@

Unlike English, code is fragile in standing the test of time.
Containers can alleviate this, as these can be created at a point
in time and then run later. Because the container is stand-alone,
all software that the code depends on remain in their current/old state.
As with hosted code, an experiment may not run with newer versions
of libraries anymore, but a containerized experiment can run regardless, 
if one refrains from updating (or only does so when there are there problems with fixing).

@@

There are some problems with providing a container with a runnable container.
First, a container is several gigabytes and need to be stored somewhere
online. 
Ideally, for Singularity, this would be at the main container hosting
site called Syslabs. Using Syslabs is free upon registration, 
yet there is a limit of 15 Gb. Also, Sylabs does not allow much metadata to
be entered, such as the URL of the article, nor the URL of hosted code.
Also, journals do not request hosted code, nor runnable code.

@@

The code that does the work in a computational experiment should
always be published together with a scholarly article.
For research to be reproducible, one ideally has access to
both the data used and the code.
In some fields, such as genetic epidemiology, the data is
sensitive, hence cannot be released,
yet there are methods being devised to run code on sensitive
data with assured privacy (\cite{zhang2016review,azencott2018machine}).

Code has additional useful information, similar to confidence intervals,
that allow a reader to gauge how much he/she trusts the results.
The most important way to determine the quality of code
is the amount of unit tests.
When following a set of best practices, such as DevOps, TDD, Agile,
writing unit tests is an essential 
part in writing code.
The amount of unit tests is an honest signal 
for code correctness (i.e. it does what it is supposed to do, as opposed
to 'it does something').
In academia, to uncover the truth, code correctness is essential.
To make a comparison with cell biologists, where working
sterile is essential to have the correct results,
unit tests allow a computational biologist to achieve the same.
